---
title: "P8105 Homework 2"
author: "Yikai Lu"
date: "`r Sys.Date()`"
output: github_document
---
# Problem 1
```{r}
library(tidyverse)

nyc_transit =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    ),
    across(route1:route11, as.character)
  )

```
#The NYC Transit dataset contains information about subway station entrances and exits, including the subway line, station name, station latitude and longitude, routes served, entry status, vending availability, entrance type, and ADA compliance. I have cleaned the data like what we did in class, I clean by cleaning the variable names, selecting only the variables needed for this problem, and converting the entry variable from a character variable (YES or NO) to a logical variable (TRUE or FALSE). After cleaning the data, the dataset contains 1,868 rows and 19 columns. 

```{r}
#How many station
nyc_transit |>
  distinct(line, station_name)
nyc_transit |>
  distinct(line, station_name) |>
  nrow()
```

#There are 465 distinct subway stations, where a station is identified by the combination of line and station_name.

```{r}
nyc_transit |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```
#There are 84 distinct ADA-compliant stations.

```{r}
# Proportion of entrances / exits without vending that allow entrance
nyc_transit |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```
#The proportion of station entrances/exits without vending that allow entrance is 0.377, or approximately 37.7%

```{r}
nyc_transit_long =
  nyc_transit |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  )
nyc_transit_long
# Number of distinct stations that serve the A train
nyc_transit_long |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

#There are 60 distinct stations that serve the A train.

```{r}
# Number of ADA-compliant stations that serve the A train
nyc_transit_long |>
  filter(
    route_name == "A",
    ada == TRUE
  ) |>
  distinct(line, station_name) |>
  nrow()

```
#Of the stations that serve the A train, 17 are ADA compliant.


#Problem 2
```{r}
library(readxl)

mr_trash =
  read_excel(
    "./data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N1170"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
  sports_balls = as.integer(round(sports_balls)),
  trash_wheel = "Mr. Trash Wheel"
  )
mr_trash
```
```{r}
professor_trash =
  read_excel(
    "./data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "Professor Trash Wheel"
  )

professor_trash
```
```{r}
gwynnda_trash =
  read_excel(
    "./data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "Gwynns Falls Trash Wheel"
  )

gwynnda_trash
```
```{r}
mr_trash =
  mr_trash |>
  mutate(year = as.numeric(year))

professor_trash =
  professor_trash |>
  mutate(year = as.numeric(year))

gwynnda_trash =
  gwynnda_trash |>
  mutate(year = as.numeric(year))

trash_wheel =
  bind_rows(
    mr_trash,
    professor_trash,
    gwynnda_trash
  )

trash_wheel
```
```{r}
trash_wheel |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  pull(weight_tons) |>
  sum(na.rm = TRUE)
trash_wheel |>
  filter(
    trash_wheel == "Gwynns Falls Trash Wheel",
    month == "June",
    year == 2022
  ) |>
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)
```
#The combined Trash Wheel dataset contains 1,188 observations and 15 variables. The data include information about each dumpster collection, such as the dumpster number, month, year, date, weight in tons, volume in cubic yards, plastic bottles, and cigarette butts. The data from Mr. Trash Wheel, Professor Trash Wheel, and Gwynns Falls Trash Wheel were imported, cleaned, and combined into one dataset. A variable was added to identify each Trash Wheel. Based on the available data, Professor Trash Wheel collected a total of 282.26 tons of trash. Gwynns Falls Trash Wheel collected 18,120 cigarette butts in June 2022.

#Problem 3
```{r}
baseline =
  read_csv(
    "./data/MCI_baseline.csv",
    skip = 1
  ) |>
  janitor::clean_names()

baseline
names(baseline)
```
```{r}
#clean data
baseline =
  baseline |>
  mutate(
    age_at_onset = na_if(age_at_onset, "."),
    age_at_onset = as.numeric(age_at_onset),
    sex = factor(
      sex,
      levels = c(0, 1),
      labels = c("Female", "Male")
    ),
    apoe4 = factor(
      apoe4,
      levels = c(0, 1),
      labels = c("Non-carrier", "Carrier")
    )
  ) |>
  filter(
    is.na(age_at_onset) | age_at_onset > current_age
  )

baseline

```
```{r}
#participants
nrow(baseline)
```
```{r}
# Number of participants
nrow(baseline)

# Number who developed MCI
baseline |>
  filter(!is.na(age_at_onset)) |>
  nrow()

# Average baseline age
baseline |>
  summarize(
    mean_age = mean(current_age, na.rm = TRUE)
  )

# Proportion of women who are APOE4 carriers
baseline |>
  filter(sex == "Female") |>
  summarize(
    proportion_apoe4 = mean(apoe4 == "Carrier", na.rm = TRUE)
  )
```
#The baseline demographic dataset was imported and cleaned by removing participants who had MCI at baseline. Sex and APOE4 carrier status were converted from numeric variables to categorical variables. After applying the inclusion criteria, the dataset contained 479 participants. Among these participants, 93 developed MCI during the follow-up period. The average baseline age was approximately 65.03 years. Among women in the study, 30% were APOE4 carriers.

```{r}
amyloid =
  read_csv(
    "./data/mci_amyloid.csv",
    skip = 1
  ) |>
  janitor::clean_names()

amyloid
names(amyloid)
```
```{r}
#Tidy the amyloid data
amyloid_long =
  amyloid_long |>
  mutate(
    amyloid_ratio = as.numeric(amyloid_ratio)
  )

amyloid_long
```

```{r}
# Rename ID variable
amyloid_long =
  amyloid_long |>
  rename(id = study_id)

# Participants only in baseline data
baseline_only =
  anti_join(
    baseline,
    amyloid_long,
    by = "id"
  )

# Participants only in amyloid data
amyloid_only =
  anti_join(
    amyloid_long,
    baseline,
    by = "id"
  )

baseline_only

amyloid_only
```
#The amyloid dataset was imported from mci_amyloid.csv. The first row was skipped because it contained descriptive labels rather than observations. The data were converted from wide to long format so that each row represents one participant at one measurement occasion. The measurement occasions are baseline, time 2, time 4, time 6, and time 8. The value "Na" was converted to a missing value (NA), and amyloid ratio was converted to a numeric variable.

```{r}
# Number of participants only in baseline data
baseline_only |>
  distinct(id) |>
  nrow()

# Number of participants only in amyloid data
amyloid_only |>
  distinct(id) |>
  nrow()
```
```{r}
# Combine baseline and amyloid data
combined_data =
  inner_join(
    baseline,
    amyloid_long,
    by = "id"
  )

combined_data

# Check dimensions
dim(combined_data)

# Number of unique participants
combined_data |>
  distinct(id) |>
  nrow()
```
```{r}
# Export the combined dataset
write_csv(
  combined_data,
  "./data/mci_combined.csv"
)
```
#There were 8 participants who appeared only in the baseline dataset and 16 participants who appeared only in the amyloid dataset. I used an inner join to retain only participants who appeared in both datasets. The resulting dataset contains 471 unique participants, 2,355 observations, and 8 variables. Each participant has up to five amyloid measurement occasions: baseline, time 2, time 4, time 6, and time 8. The combined dataset was exported as a CSV file to the data directory.


























```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

